Measuring the double-drift illusion with hand trajectories

### 2021-12-16 Manuscript.Rmd
added comparison to illusion strengths used in Heller et al, and mentioned range of K across participants in exp 2, in the caption to fig 4

---
title: "Measuring the double-drift illusion with hand trajectories"
author: 
- Bernard Marius 't Hart 1,*
- Denise Y.P. Henriques 1
- Patrick Cavanagh 1,2,3
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document:
    reference_docx: template.docx
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{amsmath}
bibliography: bibliography.bib
csl: journal-of-vision.csl
link-citations: yes
---

@@

In particular, we compare the illusion strengths we find with those reported by Heller et al. [-@Heller2021]. They show a double-drift stimulus with 36 combinations of internal and external motion for 500 ms and then have participants indicate on a ring where they perceived the path of the stimulus to intersect with this ring. While they hypothesize that illusion strength decreases with faster external motion (perhaps because of resets or limits on perception), here we will use their overall estimate of illusion strength, given by $K=0.74$.

@@

```{r fig4, echo=FALSE, message=FALSE, fig.width=6, fig.height=6, fig.cap="\\label{fig:fig4}**Figure 4:** Re-tracing Data. **A-D:** Example trajectories and reset points. All trajectory samples in transparent light blue (denser samples result in darker color). Dark red lines are, pruned, interpolated and filtered trajectories (see Methods). Purple circles indicate reset points. **A:** Trajectory with several sudden resets. **B:** 'Hit-the-wall' reset. **C:** Slow return reset. **D:** Continued accumulation of the illusion without any reset. **E:** Distribution of all 827 detected reset points. **F:** Using the angle from the origin of the trace to a point along the trajectory at half the distance between origin and reset point as a measure of illusion strength, we can compare these data to other data. For this graph, we exclude one participant who produced very few resets. The illusion is weaker than in Experiment 1 ($K=0.60$, blue line) and its strength varies between participants, but it is nevertheless clearly there."}
#plotExampleData()
plotResetPoints()
```

>>>>

---
title: "Measuring the double-drift illusion with hand trajectories"
author: 
- Bernard Marius 't Hart 1,*
- Denise Y.P. Henriques 1
- Patrick Cavanagh 1,2,3
output:
  word_document:
    reference_docx: template.docx
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{amsmath}
bibliography: bibliography.bib
csl: journal-of-vision.csl
link-citations: yes
---

@@

In particular, we compare the illusion strengths we find with those reported by Heller et al. [-@Heller2021]. They show a double-drift stimulus with 36 combinations of internal and external motion for 500 ms and then have participants indicate on a ring where they perceived the path of the stimulus to intersect with this ring. While they hypothesize that illusion strength decreases with faster external motion (perhaps because of resets or limits on perception), here we will use their overall estimate of illusion strength, given by $K=0.74$. While in _dva/s_ our stimuli have relatively high external and relatively low internal motion, compared to Heller et al. [-@Heller2021], they are within the range of tested stimuli as is the predicted illusion strength.

@@

```{r fig4, echo=FALSE, message=FALSE, fig.width=6, fig.height=6, fig.cap="\\label{fig:fig4}**Figure 4:** Re-tracing Data. **A-D:** Example trajectories and reset points. All trajectory samples in transparent light blue (denser samples result in darker color). Dark red lines are, pruned, interpolated and filtered trajectories (see Methods). Purple circles indicate reset points. **A:** Trajectory with several sudden resets. **B:** 'Hit-the-wall' reset. **C:** Slow return reset. **D:** Continued accumulation of the illusion without any reset. **E:** Distribution of all 827 detected reset points. **F:** Using the angle from the origin of the trace to a point along the trajectory at half the distance between origin and reset point as a measure of illusion strength, we can compare these data to other data. For this graph, we exclude one participant who produced very few resets. The illusion is weaker than in Experiment 1 ($K=0.60$, blue line) and its strength varies between participants (from $K=0.40$ to $K=0.81$), but it is nevertheless clearly there."}
#plotExampleData()
plotResetPoints()
```
